Air Quality Data - PM2.5 Components

### 2021-01-18 code/data-processors/air-data-processor.Rmd
update pm daily mergeing method avg of all zipcodes in 12-mile buffer

```{r load packages}
library(tidyverse)
library(here)
library(lubridate)
#library(naniar) # for replacing -999 with NA
```

@@

```{r load data}
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\state-region.RDa')
```

@@

dt.aqs.sites <- read.csv(unz(temp,"aqs_sites.csv")) %>%
  filter(
    #remove monitors closed before 2000
    year(as.Date(Site.Closed.Date)) >= 2000 | Site.Closed.Date == "",
    !State.Name %in% c("Alaska","Hawaii","Puerto Rico","Virgin Islands",
                       "Guam", "Country Of Mexico","Canada") #remove non-CONUS states
    ) %>%  
  transmute(
    Site.ID = as.factor(sprintf("%02d-%03d-%04d", #unique ID for each Site: ss-ccc-nnnn
                                as.numeric(as.character(State.Code)),
                                County.Code, Site.Number)), 
    State.Code = as.factor(sprintf("%02d", as.numeric(as.character(State.Code)))),
    County.Code = as.factor(sprintf("%03d", County.Code)),
    Latitude,
    Longitude, 
    State = state.abb[match(State.Name,state.name)],
    State = as.factor(ifelse(State.Name == "District Of Columbia","DC", State)),
    Land.Use,
    Location.Setting = as.factor(ifelse(Location.Setting == "URBAN AND CENTER CITY",
                                        "URBAN",
                                        as.character(Location.Setting)))
    ) %>% 
  merge(dt.state.region[, c("State", "Region.IV")])

@@

```{r EPA AQS Data Download}
#Using EPA.AQS.download Function
dt.spec.raw <- EPA.AQS.download (period = "daily", Index = "SPEC",
                        start = 2000, end = 2008, 
                        Code.Sym = Code.to.Sym)
save(dt.spec.raw, file = here('data', 'raw', 'spec-raw.RDa'))
#Monitor.ID: Site.ID-POC

dt.spec.raw %>% distinct(Site.ID) %>% nrow() #546
dt.spec.raw %>% 
  mutate(Year = year(Date)) %>% 
  distinct(Site.ID, Year) %>% 
  group_by(Year) %>% 
  summarise(n = n())

dt.test <- Air.Daily.Process(dt = dt.spec.raw, d = 4, m = 9, y = 4)

dt.test2 <- dt.test %>% 
  select(-Parameter.Code) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
  # for 405 cases with two methods for EC and OC (TOR & TOT), use TOT 
  mutate(EC = ifelse(!is.na(EC.TOT), EC.TOT, EC.TOR), 
         OC = ifelse(!is.na(OC.TOT), OC.TOT, OC.TOR)) %>% 
  select(-EC.TOR, -EC.TOT, -OC.TOR, -OC.TOT) %>% 
  arrange(Site.ID, POC, Date)
```

@@

dt.imprv.raw.l <- dt.imprv.raw %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter.Symbol",
               values_to = "Arithmetic.Mean",
               values_drop_na = T)

dt.test3 <- Air.Daily.Process(dt = dt.imprv.raw.l, d = 4, m = 9, y = 4)

dt.test4 <- dt.test3 %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(Arithmetic.Mean)) %>% 
    arrange(Site.ID, POC, Date)

# find the diff bw spec & improve
monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.imprv <- dt.imprv.raw %>% distinct(Monitor.ID) 
monitor.diff.imprv <- setdiff(monitor.imprv$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame() 
colnames(monitor.diff.imprv) <- 'Monitor.ID'
# improve has 32 Monitor.ID that spec does not have

site.spec <- dt.spec.w %>% distinct(Site.ID)
site.imprv <- dt.imprv.raw %>% distinct(Site.ID) 
site.diff.imprv <- setdiff(site.imprv$Site.ID, site.spec$Site.ID) %>% 
  as.data.frame()
colnames(site.diff.imprv) <- 'Site.ID'
# improve has 28 Site.ID that spec does not have
```

@@

monitor.spec <- dt.spec.w %>% distinct(Monitor.ID)
monitor.csn <- dt.csn.raw %>% distinct(Monitor.ID) 
monitor.diff.csn <- setdiff(monitor.csn$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame()
# no difference between two dataset!
# csn is not adding any new information to spec
```

@@

dt.site.carbon.method %>% group_by(Method) %>% 
  summarise(n = n())
# TOR 167
# TOT 326
# EC.TOT-OC.TOR 28
```

```{r Adding Daily PM data to pmc daily data}
load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\site-zip-closest-bz12.RDa')

# closest zipcode to each site in each year in 12-mile bufferzone
dt.site.close.zip.pmc <- site.pmc %>% 
  merge(dt.site.zip.closest.bz12) %>% 
  arrange(Site.ID, Year)

save(dt.site.close.zip.pmc, file = here('data', 'processed', 'site-close-zip-pmc.RDa'))
write.csv(dt.site.close.zip.pmc, file = here('data', 'processed', 'site-close-zip-pmc.csv'))

dt.site.close.zip.pmc %>% distinct(Site.ID) %>% nrow() #574 > 560
# 560 out of 574 sites have zipcodes in their 12-mile buffer

@@

# daily PM value for each site based on closest zipcode value
dt.pm.daily <- read.csv(here('data','processed','pm_pmc_daily.csv')) %>% 
  transmute(Site.ID = Site_ID,
            Date = mdy(chardate),
            PM)

@@

dt.pmc.pm.w %>% distinct(Site.ID) %>% nrow() # 574 > 560 > 558
dt.pmc.pm.w %>% distinct(Monitor.ID) %>% nrow()#691 > 673

@@

#sites with measurements for all components - spec & improve
dt.pmc.c <- dt.pmc.pm.w[complete.cases(dt.pmc.pm.w),] #241,624
monitor.pmc <- dt.pmc.c %>% distinct(Monitor.ID) # 691 > 673 > 515
site.pmc <- dt.pmc.c %>% distinct(Site.ID) # 574 > 558 > 471

@@

monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 691 > 673 > 515 > 331
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 558 > 471 > 318

@@

dt.pmc.sel <- merge(dt.pmc.sel, site.POC.larger) 
monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 691 > 673 > 515 > 331 > 318
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 558 > 471 > 318 > 318
dt.pmc.sel$POC <- dt.pmc.sel$Monitor.ID <- NULL
save(dt.pmc.sel, file = here('data', 'processed', 'pmc-selected.RDa'))
```

@@

load('C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\site-zip.RDa')

>>>>

```{r load packages}
library(tidyverse)
library(here)
library(lubridate)
library(naniar) # for replacing -999 with NA
```

@@

```{r load data}
dirUSSpatial <- 'C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\'
load(paste0(dirUSSpatial,'state-region.RDa'))
load(paste0(dirUSSpatial,'site-zip-closest-bz12.RDa'))
load(paste0(dirUSSpatial,'site-zip.RDa'))
```

@@

dt.aqs.sites <- read.csv(unz(temp,"aqs_sites.csv")) %>%
  filter(
    #remove monitors closed before 2000
    year(as.Date(Site.Closed.Date)) >= 2000 | Site.Closed.Date == "",
    !State.Name %in% c("Alaska","Hawaii","Puerto Rico","Virgin Islands",
                       "Guam", "Country Of Mexico","Canada") #remove non-CONUS states
    ) %>%  
  transmute(
    Site.ID = as.factor(sprintf("%02d-%03d-%04d", #unique ID for each Site: ss-ccc-nnnn
                                as.numeric(as.character(State.Code)),
                                County.Code, Site.Number)), 
    State.Code = as.factor(sprintf("%02d", as.numeric(as.character(State.Code)))),
    County.Code = as.factor(sprintf("%03d", County.Code)),
    Latitude,
    Longitude, 
    State = state.abb[match(State.Name,state.name)],
    State = as.factor(ifelse(State.Name == "District Of Columbia","DC", State)),
    Land.Use,
    Location.Setting = as.factor(ifelse(Location.Setting == "URBAN AND CENTER CITY",
                                        "URBAN",
                                        as.character(Location.Setting)))
    ) %>% 
  merge(dt.state.region[, c("State", "Region.IV")]) %>% 
  arrange(Site.ID)

@@

```{r EPA AQS Data Download}
#Using EPA.AQS.download Function
dt.spec.raw <- EPA.AQS.download (period = "daily", Index = "SPEC",
                        start = 2000, end = 2008, 
                        Code.Sym = Code.to.Sym)
save(dt.spec.raw, file = here('data', 'raw', 'spec-raw.RDa'))
#Monitor.ID: Site.ID-POC
```

```{r EPA AQS Data Exploration}
dt.spec.raw %>% distinct(Monitor.ID) %>% nrow() #659
dt.spec.raw %>% distinct(Site.ID) %>% nrow() #546

# number of sites per year
dt.spec.raw %>% 
  mutate(Year = year(Date)) %>% 
  distinct(Site.ID, Year) %>% 
  group_by(Year) %>% 
  summarise(n = n())

# number of monitors per year
dt.spec.raw %>% 
  mutate(Year = year(Date)) %>% 
  distinct(Monitor.ID, Year) %>% 
  group_by(Year) %>% 
  summarise(n = n())
```

@@

# find the diff bw spec & improve
monitor.spec <- dt.spec.w %>% distinct(Monitor.ID) #659
monitor.imprv <- dt.imprv.raw %>% distinct(Monitor.ID) #189
monitor.diff.imprv <- setdiff(monitor.imprv$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame() #32
colnames(monitor.diff.imprv) <- 'Monitor.ID'
# improve has 32 Monitor.ID that spec does not have

site.spec <- dt.spec.w %>% distinct(Site.ID) #546
site.imprv <- dt.imprv.raw %>% distinct(Site.ID) #172
site.diff.imprv <- setdiff(site.imprv$Site.ID, site.spec$Site.ID) %>% 
  as.data.frame() #28
colnames(site.diff.imprv) <- 'Site.ID'
# improve has 28 Site.ID that spec does not have
```

@@

monitor.spec <- dt.spec.w %>% distinct(Monitor.ID) #659
monitor.csn <- dt.csn.raw %>% distinct(Monitor.ID) #195
monitor.diff.csn <- setdiff(monitor.csn$Monitor.ID, monitor.spec$Monitor.ID) %>% 
  as.data.frame() #0
# no difference between two dataset!
# csn is not adding any new information to spec
```

@@

dt.site.carbon.method %>% group_by(Method) %>% 
  summarise(n = n())
# TOT 326
# TOR 167
# EC.TOT-OC.TOR 28
```

```{r Adding Daily PM data to pmc daily data}
# available site-year pmc
site.yr.pmc <- dt.pmc.w %>% 
  mutate(Year = year(Date)) %>% 
  distinct(Site.ID, Year) #3394

# all zipcodes in 12-mile bufferzone of each site each year
dt.pmc.site.all.zip.bz12  <- site.yr.pmc %>% 
  inner_join(dt.site.zip) %>% 
  arrange(Site.ID, Year, Dist) %>% 
  filter(Dist <=12) %>% 
  select(-Zip.Type, -Dist)

save(dt.pmc.site.all.zip.bz12, file = here('data', 'processed', 'pmc-site-all-zip-bz12.RDa'))
write.csv(dt.pmc.site.all.zip.bz12, file = here('data', 'processed', 'pmc-site-all-zip-bz12.csv'))

dt.pmc.site.all.zip.bz12 %>% distinct(Site.ID) %>% nrow() #574 > 558
# 558 out of 574 sites have zipcodes in their 12-mile buffer

@@

# daily PM value for each site based on closest zipcode value
dt.pm.daily <- read.csv(here('data','processed','pm_pmc_daily.csv')) %>% 
  transmute(Site.ID = Site_ID,
            Date = mdy(chardate),
            PM) %>% 
  arrange(Site.ID, Date)

@@

dt.pmc.pm.w %>% distinct(Site.ID) %>% nrow() # 574 > 558
dt.pmc.pm.w %>% distinct(Monitor.ID) %>% nrow()#691 > 673

@@

#sites with measurements for all components - spec & improve
dt.pmc.c <- dt.pmc.pm.w[complete.cases(dt.pmc.pm.w),] #243,098
monitor.pmc <- dt.pmc.c %>% distinct(Monitor.ID) # 691 > 673 > 515
site.pmc <- dt.pmc.c %>% distinct(Site.ID) # 574 > 558 > 471

@@

monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 691 > 673 > 515 > 336
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 558 > 471 > 322

@@

dt.pmc.sel <- merge(dt.pmc.sel, site.POC.larger) 
monitor.pmc <- dt.pmc.sel %>% distinct(Monitor.ID) # 691 > 673 > 515 > 336 > 322
site.pmc <- dt.pmc.sel %>% distinct(Site.ID) #574 -> 558 > 471 > 322
dt.pmc.sel$POC <- dt.pmc.sel$Monitor.ID <- NULL
save(dt.pmc.sel, file = here('data', 'processed', 'pmc-selected.RDa'))
```

@@
